### 2019-05-16 demo-doc.Rmd
add demo doc

---
title: "Workshop Demo Doc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
gss_cat
```

# Select

```{r}
gss_cat %>% 
  select(year, marital, tvhours)
```

Try it out!

Select three *different* variables

```{r}
gss_cat %>% 
  select()
```

# Filter

```{r}
gss_cat %>% 
  filter(age > 30)
```

Try it out!

Filter to include only those individuals who watch a lot of TV - more than 10 hours per *day*!

```{r}
gss_cat %>%

```

# Mutate

```{r}