### 2019-05-19 demo-doc.Rmd
update typos/edits based on presentation

---
title: "Workshop Demo Doc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Demo doc (part 1): Try it out!

Let's first load the tidyverse suite of packages.

If you haven't already, make sure to install the tidyverse, first.

*You can skip this code "chunk" if you already have the tidyverse installed; you only need to do this once*

```{r}
install.packages("tidyverse")
```

Be sure to run this next line to load the tidyverse. You do this each time you open R.

```{r}
library(tidyverse)
```

Here is the data set we will use:

```{r}
gss_cat
```

# Select

Let's select only a few variables.

```{r}
gss_cat %>% 
  select(year, tvhours)

gss_cat %>% 
  select(year, marital, tvhours)
```

**Try it out!**

Now, select three *different* variables, and run the code chunk to view the result.

```{r}
gss_cat %>% 
  select()
```

# Filter

```{r}
gss_cat %>% 
  filter(age > 60)
```

**Try it out!**

Filter to include only those individuals who watch a lot of TV - more than 10 hours per *day*!

Be sure to run the result to see whether it did what you think it should do!

```{r}
gss_cat %>%

```

# Mutate

Let's create a new variable for those 60 years of age or older.

```{r}
gss_cat %>% 
  mutate(potential_senior = if_else(age >= 63.5, 1, 0),
         super_senior = if_else(potential_senior == 1 & tvhours < 2, 1, 0)) %>% 
  count(potential_senior, super_senior)
```

Let's save the results back to `gss_cat` and make sure that it worked.

```{r}
gss_cat <- gss_cat %>% 
  mutate(potential_senior = ifelse(age >= 63.5, TRUE, FALSE))
```

Adding another variable for folks who are middle-aged.

```{r}
gss_cat <- gss_cat %>% 
  mutate(middle_aged = ifelse(age >= 30 & age <= 40, "yes", "no"))
```

**Try it out!**

Can you create a new variable based upon either the *year* or *tvhours* variable using `if_else()`?

```{r}
gss_cat %>% 
  mutate()
```

Then, save the results back to `gss_cat` (check out the code two chunks above for an example!).

```{r}
gss_cat %>% 
  mutate()
```

# Demo doc (part 2): Try it out!

`geom_jitter()` is exactly like geom_point, but with a small amount of x- and y-axis random noise added.

```{r}
ggplot(gss_cat, aes(x = age, y = tvhours)) +
  geom_jitter()
```

**Try it out!**

Can you add a line representing the linear relationship between the two varibles? Add another layer using a `geom_` function.

```{r}
# hint: check out how a line was added in the presentation?

ggplot(gss_cat, aes(x = age, y = tvhours)) +
  geom_jitter() +
  geom_
```

Let's try a different theme.

```{r}
ggplot(gss_cat, aes(x = age, y = tvhours)) +
  geom_jitter() +
  theme_bw()
```

**Try it out!**

Can you add a different theme? Start typing `theme_` and see what shows up! Be sure to run the code to see the results!

```{r}
ggplot(gss_cat, aes(x = age, y = tvhours)) +
  geom_jitter() +
  theme_
```

Does the relationship look to be different by year?

```{r}
ggplot(gss_cat, aes(x = age, y = tvhours)) +
  geom_jitter() +
  facet_wrap(~year)
```

**Try it out!**

Can you add lines of best fit to the above plots?

```{r}
ggplot(gss_cat, aes(x = age, y = tvhours)) +
  geom_jitter() +
  geom_
  facet_wrap(~year)
```

# Demo doc (part 3): Try it out!

Let's run a regression.

```{r}
m1 <- lm(tvhours ~ age, data = gss_cat)
summary(m1)
```

**Try it out!**

Can you add another variable to the model? If you need to check what variables are in `gss_cat`, print that object to the console.

```{r}
m2 <- lm(tvhours ~ age + , data = gss_cat)
summary(m1)
```